---
title: "Programming Assignment 4"
subtitle: "Data Science"
author: "Andrea Valeria Carbone"
date: "2025/03/25 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r}
library(tidyverse)
library(dplyr)
library(ds4ling)
library("here")
```
---

```{r}
data <- read.csv(here("data_raw", "fricatives", "fricatives.csv"))
```
---
```{r}
#data tidying

data_tidy <- data |>
  pivot_longer(cols = c(s_cog, sh_cog),
               names_to = "phoneme",
               values_to = "cog")|>
pivot_longer(cols = c(s_skewness, sh_skewness),
             names_to = "phoneme_skewness",
             values_to = "skewness")|>
  select("phoneme", "cog", "skewness")|>
  mutate(phoneme = gsub("_cog", "", phoneme))
         
        
         
#table of descriptive statistics
         summary(data_tidy)
         print(summary(data_tidy))
         write.csv(data_tidy, "../data_tidy/data_tidy.csv")
```
---
```{r}
 print(data_tidy)
```
---
#Table of descriptive statistics
```{r}
summary_table <- data_tidy|>
  group_by(phoneme)|>
  summarise(
    mean_skew = mean (skewness),
    mean_cog = mean (cog),
    mean_skew = median (skewness),
    mead_cog = median (cog)
    
  )
knitr::kable(summary_table)
```
---
#Tidy data
```{r}
write.csv(data_tidy, "../data_tidy/data_tidy.csv")
```
---
#Boxpot of cog as a function of phoneme

```{r}
data_tidy|>
  ggplot() +
  (aes(x = phoneme, y = cog, fill = phoneme)) +
  geom_boxplot() +
  labs(title = "cog",
       x = "phoneme", y = "cog")
       

```

---
#skewness as a function of phoneme

```{r}
ggplot(data_tidy,
       aes(x = phoneme, y = skewness, fill = phoneme)) +
  stat_summary(fun = mean) +
  labs(title = "skewness as a function of phoneme",
       x = "phoneme", y = "skewness")
       
```

#cog as a function of skewness for the s segments
---
```{r}
data_s <- data_tidy |> 
  
  filter(phoneme == "s")

model <- lm(
  cog ~ skewness,
  data = data_s)

print(model)
```
---
#scatterplot
```{r}
data_tidy |>
ggplot() +
  aes(x = cog, y = skewness) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "gravity as a function of skewness",
      x = "cog", y = "skewness")

```

---
#Table model
```{r}
knitr::kable(summary(model)$coef)
```


---
#Model diagnostics
```{r}

diagnosis(model)
```
---
# report
There is a positive correlation between skewness and center of gravity, so as skewness increases, the center of gravity also shifts in the same direction, and the increase of COG is statistically significant with pi values around 0. 



---
```{r}
data_2 <- read_csv(here("data_raw", "fricatives", "assumptions.csv"))
```

```{r}
data_2|>
  ggplot() +
  geom_point() +
  aes(x = x, y = y) +
  labs (x ="x", y ="y", tile = "Assumptions")

```
---
#explanation

There is not real data, but a dinosaur.

---









